\begin{abstract}
Large language models have fundamentally transformed software development, enabling a new generation of AI-powered coding tools that range from inline code completion to fully autonomous software agents. The rapid proliferation of these tools---including GitHub Copilot, ChatGPT, Claude, Cursor, Devin, and SWE-agent---has created a fragmented landscape where researchers and practitioners struggle to understand capabilities, compare systems, and identify research opportunities.

This survey provides a comprehensive overview of AI coding agents from the LLM era (2022--present). We present a taxonomy that categorizes systems along three dimensions: autonomy level (from reactive assistance to autonomous execution), interaction scope (from single-line completion to repository-wide changes), and integration mode (IDE extensions, chat interfaces, CLI tools, and standalone platforms). Using this framework, we systematically analyze four categories of tools: code completion systems, conversational coding assistants, autonomous coding agents, and AI-integrated development environments.

We examine evaluation benchmarks including HumanEval, MBPP, and SWE-bench, discussing their strengths and limitations for measuring real-world utility. Finally, we identify open challenges in reliability, security, code quality, and human-agent collaboration, outlining a research agenda for the software engineering community. Our survey aims to help researchers navigate this rapidly evolving field and identify high-impact research directions.
\end{abstract}
